---
title: "Analysis: Google Search Interest in Immigration-Related Topics vs Electoral Swing Per County"
author: "Dowland Aiello, Roy Lin, Winston Qi"
date: "2025-06-08"
output: html_document
---

```{r setup ,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Overview

Herein, we explore the relationship between google search query frequency for topics related to immigration and the electoral swing in the Presidential election at a given point in time. We are particularly interested in the relationship between the change in search frequency and the magnitude of the electoral swinng.

## 2. Data Cleaning

See the `scripts/` folder for utilities we wrote to clean our data. Notably, Google trends data for city / region over varying time frames were combined from multiple CSV files into a single table, which looks like this:

```{r}
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)

trendsdf <- read.csv("../../data_cleaned/google_trends/summarized.csv")
trendsdf %>% head(5)
```

## 3. Electoral Swing from 2004 - 2008 Presidential Elections

### 3.1 Data Overview

Here is an overview of the electiondf dataset:

```{r}
electiondf <- read.csv("../../data_cleaned/harvard_election/countypres_2000-2020.csv")
electiondf %>% sample_n(10)
```

### 3.2 Data Cleaning

For this particular analysis, we are interested in the swing from 2004 to 2008. We can get all data from these date ranges like such:

```{r}
electiondf %>%
    filter(year == 2004 | year == 2008) %>%
    sample_n(5)
```

We can gather all votes in 2004 and 2008 per county via the pivot function:

```{r}
votestimeframe <- electiondf %>%
  select(year, state, county_name, party, office, candidatevotes, totalvotes) %>%
  # Select only republican so that further analysis can have red/blue as meaningful colors
  filter(office == "US PRESIDENT") %>%
  # Collapse votes in year n, n1 into columns n, n1
  group_by(year, state, county_name, office, party) %>%
  summarise(year, state, county_name, candidatevotes) %>%
  ungroup() %>%
  distinct(party, state, county_name, year, .keep_all = TRUE) %>%
  select(party, state, county_name, year, candidatevotes) %>%
  pivot_wider(id_cols = c(state, county_name), names_from = c(party, year), values_from = c(candidatevotes))

votestimeframe %>% head(10)
```

### 3.3 Swing Calculation

We can then calculate the swing like such:

```{r}
votestimeframe %>%
    group_by(state, county_name) %>%
    summarise(swing = .data[["REPUBLICAN_2008"]] - .data[["REPUBLICAN_2004"]]) %>%
    head(10)
```

## 4. Change in Google Trends Over a Timeframe

### 4.1 Change Overview

Now, we calculate the change in average query incidence from one timeframe to the next. We use a similar pivoting method to get (date ranges, queyr_incidence) as columns.

```{r}
trendsdf %>%
  pivot_wider(id_cols = c(Region, City, DMA), names_from = daterange, values_from = query_incidence, values_fn = max) %>%
  (\(df) df %>% mutate("change_04-12" = df[["08-12"]] - df[["04-08"]])) %>%
  head(10)
```

### 4. Generic Function

Now, we make a reusable function which calculates the change over a given time period for all regions.

```{r}
change_time_period <- function(date1, date2) {
  trendsdf %>%
  pivot_wider(id_cols = c(Region, City, DMA), names_from = daterange, values_from = query_incidence, values_fn = max) %>%
  (\(df) df %>% mutate("change_query_incidence" = df[[date1]] - df[[date2]]))
}

change_time_period("08-12", "04-08") %>% head(10)
```

## 5. Heatmap: Query Incidence Swing vs Electoral Swing (2004 election to 2012 election)

### 5.1 Data Overview

We first match cities from Google trends to electoral data. We do so by checking for cities in the election data which are substrings of the Google trends DMA column.
Note that the line of code for doing a fuzzy merge uses the fuzzyjoin package.

```{r}
library(fuzzyjoin)

interest_swing_08_12 <- change_time_period("08-12", "04-08") %>% mutate(county = toupper(DMA)) %>% select(county, change_query_incidence)
swing_county <- votestimeframe %>%
    group_by(state, county_name) %>%
    summarise(swing = abs(.data[["REPUBLICAN_2012"]] / (.data[["REPUBLICAN_2012"]] + .data[["DEMOCRAT_2012"]]) - .data[["REPUBLICAN_2004"]] / (.data[["REPUBLICAN_2004"]] + .data[["DEMOCRAT_2004"]]))) %>%
    mutate(county = county_name) %>%
    select(county, swing)

regex_inner_join(interest_swing_08_12, swing_county, by = "county") %>% head(10)
```

Looks good.

### 5.2 The Heatmap

To generate a heatmap, we get latitutde and longitudes for each county, then calculate the local slope of the linear regression between query incidence change and electoral swing for each city.

```{r}
library(ggmap)

#register_google(key = Sys.getenv("GOOGLE_API_KEY"), write = FALSE)

#with_lat_long <- regex_inner_join(interest_swing_08_12, swing_county, by = "county") %>%
#    select(county = county.y, swing, change_query_incidence, state) %>%
#    (\(df) df %>% mutate(lat_long = geocode(output = "latlon", location = paste(df$county, ", ", df$state), method = "census")))

#with_lat_long %>% write.csv("../../data_cleaned/swing_interest_vs_electoral_04_12_lat_long.csv")
#with_lat_long %>% head(10)

with_lat_long <- read.csv("../../data_cleaned/swing_interest_vs_electoral_04_12_lat_long.csv") %>%
    mutate(lat = lat_long.lat, long = lat_long.lon) %>%
    arrange(lat, long)
with_corr <- with_lat_long %>%
    filter(!is.na(change_query_incidence) & !is.na(swing))

with_corr %>% select(change_query_incidence, swing, lat, long) %>% head(10)
```

Looks good.
Now, we generate the heatmap.

```{r}
library(maps)
library(ggplot2)
library(scales)

plot <- ggplot(data = map_data("state") %>% filter(long > -140)) +
  geom_polygon(aes(x = long, y = lat, group = group), color = "white") +
  geom_point(data = with_corr %>% filter(long > -140), aes(x = long, y = lat, color = as.numeric(swing), size = as.numeric(change_query_incidence))) +
  scale_colour_gradient2(low = "blue", high = "red", mid = "gray", midpoint = 0, limits=c(-0.5, 0.5)) +
  labs(
    title = "Change in Search Interest for \"Immigration\" vs Electoral Swing from 2004 - 2012",
    color = "Swing in Election Results (Red = Became More Republican)",
    size = "Change in Search Interest for \"Immigration\"",
  ) +
  scale_size_binned(n.breaks = 5, range = c(0.25, 2.5), limits = c(-50, 100)) +
  theme(text = element_text(size = 8), legend.position = "bottom")

ggsave("election_04_to_2012_heat_map_query_change_election_result.png", bg = "white")

plot
```


### 5.3 Comparison to 2020 - 2024 change

Here, we generate the same heatmap, but for the 2020 - 2024 election.


```{r}
library(fuzzyjoin)

interest_swing_08_12_16_20 <- change_time_period("08-12", "16-20") %>% mutate(county = toupper(DMA)) %>% select(county, change_query_incidence)
swing_county <- votestimeframe %>%
    group_by(state, county_name) %>%
    summarise(swing = abs(.data[["REPUBLICAN_2016"]] / (.data[["REPUBLICAN_2016"]] + .data[["DEMOCRAT_2016"]]) - .data[["REPUBLICAN_2008"]] / (.data[["REPUBLICAN_2008"]] + .data[["DEMOCRAT_2008"]]))) %>%
    mutate(county = county_name) %>%
    select(county, swing)

regex_inner_join(interest_swing_08_12_16_20, swing_county, by = "county") %>% head(10)
```

Now, we geocode the data.

```{r}
#library(ggmap)

#register_google(key = Sys.getenv("GOOGLE_API_KEY"), write = FALSE
#)
#with_lat_long <- regex_inner_join(interest_swing_08_12_16_20, swing_county, by = "county") %>%
#    select(county = county.y, swing, change_query_incidence, state) %>%
#    (\(df) df %>% mutate(lat_long = geocode(output = "latlon", location = paste(df$county, ", ", df$state), method = "census")))

#with_lat_long %>% write.csv("../../data_cleaned/swing_interest_vs_electoral_08_20_lat_long.csv")
#with_lat_long %>% head(10)
```

We cache this data for convenience.

```{r}
library(maps)
library(ggplot2)
library(purrr)
library(scales)

with_lat_long.2 <- read.csv("../../data_cleaned/swing_interest_vs_electoral_08_20_lat_long.csv") %>%
    mutate(lat = lat_long.lat, long = lat_long.lon) %>%
    arrange(lat, long)
with_corr.2 <- with_lat_long.2 %>%
    filter(!is.na(change_query_incidence) & !is.na(swing))

with_corr.2 %>% select(change_query_incidence, swing, lat, long) %>% head(10)

plot <- ggplot(data = map_data("state") %>% filter(long > -140)) +
  geom_polygon(aes(x = long, y = lat, group = group), color = "white") +
  geom_point(data = with_corr.2 %>% filter(long > -140), aes(x = long, y = lat, color = as.numeric(swing), size = as.numeric(change_query_incidence))) +
  scale_colour_gradient2(low = "blue", high = "red", mid = "gray", midpoint = 0, limits=c(-0.5, 0.5)) +
  labs(
    title = "Change in Search Interest for \"Immigration\" vs Electoral Swing from 2008 - 2016",
    color = "Swing in Election Results (Red = Became More Republican)",
    size = "Change in Search Interest for \"Immigration\"",
  ) +
  scale_size_binned(n.breaks = 5,range = c(0.25, 2.5), limits = c(-50, 100)) +
  theme(text = element_text(size = 8), legend.position = "bottom")


ggsave("election_08_to_2020_heat_map_query_change_election_result.png", bg = "white")

plot
```
