---
title: "Analysis: Google Search Interest in Immigration-Related Topics vs Electoral Swing Per County"
author: "Dowland Aiello, Roy Lin, Winston Qi"
date: "2025-06-08"
output: html_document
---

```{r setup ,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Overview

Herein, we explore the relationship between google search query frequency for topics related to immigration and the electoral swing in the Presidential election at a given point in time. We are particularly interested in the relationship between the change in search frequency and the magnitude of the electoral swinng.

## 2. Data Cleaning

### 2.1 Region / city name sample

#### 2.1.1 Google trends dat

```{r, echo = TRUE}
library(dplyr)
library(purrr)

cities_immigration_search_df <- read.delim("./data_cleaned/google_trends/immigration_query_cities.csv")
summary(cities_immigration_search_df)
```

## 3. Results

### 3.1. 2004 United States presidential election

```{r, echo = TRUE}

```

### 3.2. 2008 United States presidential election

```{r, echo = TRUE}

```
