---
title: "electoral_deportation_results"
author: "Winston Qi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup ,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Overview
How are presidential election results correlated with immigration policy (measured via deporations and encounters at the border)? - Winston

DO: 
- Change in votes vs deportations 
- Change in votes vs encounters
- Change in border encounters vs deportations
- apprehensions data set (encounters) vs removals (deportations)
- Create variable border to determine border vs. nonborder states (make variable through mutate)
- sample from aggregate data
- Port of departure? (from deportation datasets)

```{r}
library(tidyverse)
dp_df1 <- read_delim("../../data_cleaned/removals/2023-ICFO_42034_Removals_FY13-12_LESA-STU_FINAL- Redacted_raw.xlsx.csv.gz_1.csv.gz")
#View(deport_ds1)
dp_df2 <- read_delim("../../data_cleaned/removals/2023-ICFO_42034_Removals_FY13-12_LESA-STU_FINAL- Redacted_raw.xlsx.csv.gz_2.csv.gz")
#View(deport_ds2)
dp_df3 <- read_delim("../../data_cleaned/removals/2023-ICFO_42034_Removals_FY13-12_LESA-STU_FINAL- Redacted_raw.xlsx.csv.gz_3.csv.gz")

app_df1 <- read_delim("../../data_cleaned/apprehensions_origins/Family Units apprehended along the SWB FY2000 Redacted_raw.xlsx.csv.gz_1.csv.gz")

election_results <- read_delim("../../data_cleaned/harvard_election/countypres_2000-2020.csv")
#View(election_results)
```
```{r}
ice_rem <- read_delim("../../data_cleaned/removals/ICE_Removals_FY2012-FY2014-REDACTED_raw.xlsx.csv.gz_1.csv.gz")
```

## 2 Election results 2000-2020 

```{r}
election_results %>% 
  select(year, party, candidatevotes) %>% 
  summarize(year, party, candidatevotes) %>% 
  sample_n(10) 
  # 
  # how many states voted for each party?
  # ggplot(aes(year, candidatevotes, fill = party)) +
  # geom_line()

# mutate(tot_demvotes = sum(), tot_repvotes = sum(), tot_othervotes = sum())
#

election_results %>% 
  select(year, party, candidatevotes) %>% 
  summarize(year, party, candidatevotes = sum(candidatevotes)) %>% 
  #  summarize(year, party %in% c("REPUBLICAN", "DEMOCRAT", "OTHER"), candidatevotes = sum(candidatevotes)) %>% 
  #  summarize(party$"OTHER" == sum(party != "DEMOCRAT" || party != "REPUBLICAN")) %>% 
  ggplot(aes(x = year, y = candidatevotes, fill = party)) +
  geom_col()

# election_results %>% 
#   select(year, party) %>% 
#   ggplot(aes(x = year, )) +
#   geom_bar()

```

```{r}
votestimeframe <- election_results %>% 
  select(year, state, county_name, party, office, candidatevotes, totalvotes) %>%
  # select only these cat. variables
  filter(office == "US PRESIDENT") %>% 
  group_by(year, state, county_name, party, office) %>% 
  summarise(year, state, county_name, candidatevotes) %>%
  ungroup() %>%
  distinct(party, state, county_name, year, .keep_all = TRUE) %>%
  select(party, state, county_name, year, candidatevotes) %>%
  pivot_wider(id_cols = c(state, county_name), names_from = c(party, year), values_from = c(candidatevotes))
  
votestimeframe %>% 
  head(10)
  #sample_n(10)
```


Separate data set including border variable (?):

```{r}
# include votestimeframe dataframe so when running code on rmarkdown console
# it is called for the clean run and won't give an error 
# nvm doesn't work

# filter works finally
bordervotes <- votestimeframe %>% 
  filter(state %in% c("CALIFORNIA", "ARIZONA", "NEW MEXICO", "TEXAS")) %>% 
  sample_n(10)
bordervotes
  

# works yessss
border_status <- votestimeframe %>% 
  mutate(border = state %in%
           c("CALIFORNIA", "ARIZONA",
             "NEW MEXICO", "TEXAS"))
border_status %>% 
    sample_n(10)



# works-ish (all states are non-border ATM), unique(border_status$border) only returns non-border"
# border_status <- votestimeframe %>% 
#   mutate(border = ifelse(state == state %in%
#                            c("CALIFORNIA", "ARIZONA",
#                              "NEW MEXICO", "TEXAS"),
#                          "border", "non-border")) %>% 
#   sample_n(10)
# border_status

  # str_detect(c("CALIFORNIA", "ARIZONA", "NEW MEXICO", "TEXAS")) %>% 
  # head(5)

# borderstates = c("CALIFORNIA", "ARIZONA", "NEW MEXICO", "TEXAS")
# str_detect(votestimeframe$state, borderstates)

#   mutate(border_state = ifelse(state == c("CALIFORNIA", "ARIZONA", "NEW MEXICO", "TEXAS"), "border", "non-border")) %>%
#   sample_n(10)

# California, Arizona, New Mexico, Texas
# class(votestimeframe$state)
```


Same data manipulation with deportations and border encounters:

FIND # OF DEPORTATIONS, APPREHENSIONS, & ENCOUNTERS BY YEAR
CORRESPOND IT WITH ELECTION RESULTS
```{r}
# dp_df1 %>% 
#   filter(na.rm = TRUE) %>% 
#   dim()
# sample_n()   ?


# getting somewhere but not sure how ATM
border_dp <- dp_df1 %>% 
  select(contains(c("CA", "AZ", "TX"))) %>% 
  #mutate(border = ...3 %in% contains(c("CA", "AZ", "TX"))) %>% 
  # sample_n(10)
  head(10)
border_dp
  #summarize()


# break up into dates, specifically want years
test_df <- read_delim("USBP Apprehensions FY2012_FORRELEASE_raw.csv")








ice_rem_raw <- read_delim("ICE_Removals_FY2012-FY2014-REDACTED_raw.csv")
# View(ice_rem_raw)
# ok so the deportation data was kinda messed up
# View(test_df)
totaldep_by_year <- test_df %>% 
  select(...8) %>% 
  mutate(year = ...8 %in% c())

ice_rem_raw2 <- read_delim("2023-ICFO_42034_Removals_FY13-12_LESA-STU_FINAL- Redacted_raw.csv")
View(ice_rem_raw2)

ice_rem_raw2015 <- read_delim("ICE_Removals_FY2015-FY2023_YTD-REDACTED_raw.csv")
View(ice_rem_raw2015)

ice_rem_raw16_14 <- read_delim("2023-ICFO_42034_Removals_FY16-14_LESA-STU_FINAL_Redacted_raw.csv")
View(ice_rem_raw16_14)

ice_rem_raw19_17 <- read_delim("2023-ICFO_42034_Removals_FY19-17_LESA-STU_FINAL_Redacted_raw.csv")
View(ice_rem_raw19_17)

ice_rem_raw6 <- read_delim("")


# date_sep <- function(data, date_col) {
#   mutate(data, date = as_date(date_col, format = "%m/%d/%Y")) %>%
#   mutate(year = as.numeric(format(date, format = "%Y")),
#          month = as.numeric(format(date, format = "%m")),
#          day = as.numeric(format(date, format = "%d")))
# }
# 
# 
# 
# date_sep(ice_rem_raw, ice_rem_raw$...2)
# 
# totalice_yr11_12_test <- ice_rem_raw %>% 
#   select(...2) %>% 
#   date_sep(,...2) %>% 
#   select(year, month, day, date)
# 
# totalice_yr11_12_test

# works
# shows # of deportations in current given years
totalice_yr11_12 <- ice_rem_raw %>% 
  select(...2) %>%  # , ...3
  #as.Date() %>% 
  #mutate(date = as.Date(...2, "%m/%d/%y")) %>% 
  mutate(date = as_date(...2, format = "%m/%d/%Y")) %>% 
  # finally got a working date variable
  mutate(year = as.numeric(format(date, format = "%Y")),
         month = as.numeric(format(date, format = "%m")),
         day = as.numeric(format(date, format = "%d"))) %>% 
  select(year, month, day, date)
  #mutate(deports = nrow()) %>% 
  # ok now working year, month, & day variables too

totalice_yr11_12
# function




totalice_yr13 <- ice_rem_raw2 %>% 
  select(`ERO-LESA Statistical Tracking Unit`) %>% 
  mutate(date = as_date(`ERO-LESA Statistical Tracking Unit`, format = "%m/%d/%Y")) %>% 
  mutate(year = as.numeric(format(date, format = "%Y")),
         month = as.numeric(format(date, format = "%m")),
         day = as.numeric(format(date, format = "%d"))) %>% 
  select(year, month, day, date)

totalice_yr13


totalice_yr14_15 <- ice_rem_raw2015 %>% 
  select(...3) %>% 
  mutate(date = as_date(...3, format = "%m/%d/%Y")) %>% 
  mutate(year = as.numeric(format(date, format = "%Y")),
         month = as.numeric(format(date, format = "%m")),
         day = as.numeric(format(date, format = "%d"))) %>% 
  select(year, month, day, date)
 
totalice_yr16_14 <- ice_rem_raw16_14 %>% 
  select(`ERO-LESA Statistical Tracking Unit`) %>% 
  mutate(date = as_date(`ERO-LESA Statistical Tracking Unit`, format = "%m/%d/%Y")) %>% 
  mutate(year = as.numeric(format(date, format = "%Y")),
         month = as.numeric(format(date, format = "%m")),
         day = as.numeric(format(date, format = "%d"))) %>% 
  select(year, month, day, date)

totalice_yr16_14

totalice_yr19_17 <- ice_rem_raw19_17 %>% 
  select(`ERO-LESA Statistical Tracking Unit`) %>% 
  mutate(date = as_date(`ERO-LESA Statistical Tracking Unit`, format = "%m/%d/%Y")) %>% 
  mutate(year = as.numeric(format(date, format = "%Y")),
         month = as.numeric(format(date, format = "%m")),
         day = as.numeric(format(date, format = "%d"))) %>% 
  select(year, month, day, date)

totalice_yr19_17


totalice_yr11_12
totalice_yr13
totalice_yr14_15
totalice_yr16_14
totalice_yr19_17


totalice_by_year <- rbind(totalice_yr11_12, totalice_yr13, totalice_yr14_15, totalice_yr16_14, totalice_yr19_17) %>% 
  select(date, year, month, day)
  # NEED TO SELECT B4 MERGING DATASETS

totalice_by_year
sample_n(totalice_by_year, 20)

# turn all into departed date? or simply discard extra columns

# works
icedeports_by_yr <- totalice_by_year %>% 
  select(year) %>% 
  na.omit() %>% 
  count(year)

# finally got the correct function
# %>% 
#   mutate(n = as.double(icedeports_by_yr$n))
  # this works but technically dont need bc change the variable type to double in pivot_longer

  # thought this would work   
  # mutate(deportations = count(year))
  
colnames(icedeports_by_yr)[colnames(icedeports_by_yr) == "n"] <- "Deportations"
icedeports_by_yr

# icedeports_by_yr %>% 
#   mutate(n = as.double(icedeports_by_yr$n))
# ok conversion to double worked?
# yeah just need to store it in a variable

#as.double(icedeports_by_yr$n)

#icedeports_by_yr

# why doesnt this work
# Figured it out: use count()
# unique(totalice_by_year$year) 
#summarize(n = nrow())
# unique(nrow(icedeports_by_yr))



# get # of distinct years, then count how many deportations in each year
# nrow(totalice_by_year$year == 2010)
# 
# ice_rem_raw %>% 
#   pivot_wider() %>% 
#   head(10)
# unique(ice_rem_raw$`ERO-LESA Statistical Tracking Unit`)
  # can also sort by border status


# app_df1 %>% 
  
  #pivot longer? do it by year
# check na.rm

```

Create plots with variables

barplot?

line/scatterplot of election results?

properly format deportation dates, check dim aka how many occur between every election, if more occur during certain years perhaps a pattern



line graph:
x axis = time, y = count, fill/color = votes (dem/rep), deportations, encounters etc.
graph difference in # of dem/rep votes?

bar graph (election results):
x axis (categories) = year, y = # of votes

## 3 Deportations vs Election Results

```{r}
dem_votes <- election_results %>% 
  select(year, party, candidatevotes) %>% 
  filter(party %in% "DEMOCRAT") %>% 
  group_by(year, party) %>% 
  # think this worked? but # of total votes seems a bit 
  # actually nvm that seems accurate
  summarize(tot_votes = sum(candidatevotes))
  #currently have total aggregate votes for democrats i think, now need to separate by year
dem_votes

rep_votes <- election_results %>% 
  select(year, party, candidatevotes) %>% 
  filter(party %in% "REPUBLICAN") %>% 
  group_by(year, party) %>% 
  summarize(tot_votes = sum(candidatevotes))
rep_votes

total_votes_by_yr <- rbind(dem_votes, rep_votes)
# data.frame(dem_votes, rep_votes) kinda works
# rbind is a bit better
total_votes_by_yr

# maybe pivot to wide or long form of data? so dem votes and rep votes r 2 columns
wide_votes <- total_votes_by_yr %>% 
  pivot_wider(id_cols = year, names_from = party, values_from = tot_votes)

wide_votes
# wow that worked

dep_vs_votes <- rbind(wide_votes, icedeports_by_yr) %>% 
  arrange(year)
dep_vs_votes

long_df <- dep_vs_votes %>% 
    pivot_longer(cols = c("DEMOCRAT", "REPUBLICAN", "Deportations"),
                 names_to = "Number of:", names_transform = list(n = as.double),
                 values_to = "number") %>% 
  na.omit()

long_df


# put dem, rep, & n variables into one variable specifying variable type? e.g. dem_votes, rep_votes, deportations


# total_votes_by_yr <- election_results %>% 
#   select(year, party, candidatevotes) %>% 
#   filter(party %in% c("DEMOCRAT", "REPUBLICAN")) %>% 
#   mutate(tot_dem = sum())
#   count(candidatevotes)
# find out total # of votes for each party for each year
# check votestimeframe
# sum()

# str_replace(icedeports_by_yr[2,]) = "deportations"
# icedeports_by_yr
  
ggplot(long_df, aes(x = year, y = number, col = `Number of:`)) +
  geom_point() +
  geom_line() +
  labs(title = "Deportations and votes, by year")
# add difference btwn dem and rep votes as a variable

long_df %>% 
  filter(`Number of:` == "Deportations") %>% 
  ggplot(aes(x = year, y = number)) +
  geom_point() +
  geom_line() +
  labs(title = "Deportations by year")
```

