---
title: "electoral_deportation_results"
author: "Winston Qi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup ,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Overview
How are presidential election results correlated with immigration policy (measured via deportations and apprehensions at the border)? - Winston

DO: 
DONE - Change in votes vs deportations 
- Change in votes vs encounters
- Change in border encounters vs deportations
- apprehensions data set (encounters) vs removals (deportations)
- Create variable border to determine border vs. nonborder states (make variable through mutate)
- sample from aggregate data (dont need bc have full data)
- Port of departure? (from deportation datasets)

```{r}
# install.packages("tidyverse")
library(tidyverse)
```

```{r, echo = FALSE}
# DEPORTATION DATASETS
# ice_rem <- read_delim("../../data_cleaned/removals/ICE_Removals_FY2012-FY2014-REDACTED_raw.xlsx.csv.gz_1.csv.gz")

# test_df <- read_delim("USBP Apprehensions FY2012_FORRELEASE_raw.csv")

ice_rem_raw <- read_delim("deportations_data/ICE_Removals_FY2012-FY2014-REDACTED_raw.csv")

ice_rem_raw2 <- read_delim("deportations_data/2023-ICFO_42034_Removals_FY13-12_LESA-STU_FINAL- Redacted_raw.csv")

ice_rem_raw2015 <- read_delim("deportations_data/ICE_Removals_FY2015-FY2023_YTD-REDACTED_raw.csv")

ice_rem_raw16_14 <- read_delim("deportations_data/2023-ICFO_42034_Removals_FY16-14_LESA-STU_FINAL_Redacted_raw.csv")

ice_rem_raw19_17 <- read_delim("deportations_data/2023-ICFO_42034_Removals_FY19-17_LESA-STU_FINAL_Redacted_raw.csv")

ice_rem_raw23_20 <- read_delim("deportations_data/2023-ICFO_42034_Removals_FY23-20_LESA-STU_FINAL_Redacted_raw.csv")
```

```{r}
# APPREHENSIONS DATASETS
app_2000 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2000 Redacted_raw.csv")
# View(app_2000)
app_2001 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2001 Redacted_raw.csv")
# View(app_2001)
app_2002 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2002 Redacted_raw.csv")
app_2003 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2003 Redacted_raw.csv")
app_2004 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2004 Redacted_raw.csv")
app_2005 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2005 Redacted_raw.csv")
app_2006 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2006 Redacted_raw.csv")

app_2007 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2007 Redacted_raw.csv")
app_2008 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2008 Redacted_raw.csv")
app_2009 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2009 Redacted_raw.csv")
app_2010 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2010 Redacted_raw.csv")
app_2011 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2011 Redacted_raw.csv")
app_2012 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2012 Redacted_raw.csv")
app_2013 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2013 Redacted_raw.csv")
app_2014 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2014 Redacted_raw.csv")
app_2015 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2015 Redacted_raw.csv")
app_2016 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2016 Redacted_raw.csv")
app_2017 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2017 Redacted_raw.csv")
app_2018 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2018 Redacted_raw.csv")
app_2019 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2019 Redacted_raw.csv")
app_2020 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2020 Redacted_raw.csv")
app_2021 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2021 Redacted_raw.csv")
app_2022 <- read_delim("apprehensions_data/Family Units apprehended along the SWB FY2022 Redacted_raw.csv")
```


```{r}
app_00_06 <- rbind(app_2000, app_2001, app_2002 , app_2003, app_2004, app_2005, app_2006) 
app_00_06 <- app_00_06 %>% 
  select(`U.S. Border Patrol Nationwide Apprehensions `) %>% 
  filter(str_detect(`U.S. Border Patrol Nationwide Apprehensions `, "FY2"))
# app_00_06

# app_07_15 <- data.frame()
# for(i in 7:15) {
#   app_07_15 <- rbind("app_200" + i)
# }

app_07_15 <- rbind(app_2007, app_2008, app_2009, app_2010, app_2011, app_2012, app_2013, app_2014, app_2015)
app_07_15 <- app_07_15 %>% 
  select(`U.S. Border Patrol Nationwide Apprehensions `) %>% 
  filter(str_detect(`U.S. Border Patrol Nationwide Apprehensions `, "FY2"))

app_16_22 <- rbind(app_2016, app_2017, app_2018, app_2019, app_2020, app_2021, app_2022)
app_16_22 <- app_16_22 %>% 
  select(`U.S. Border Patrol Nationwide Apprehensions `) %>% 
  filter(str_detect(`U.S. Border Patrol Nationwide Apprehensions `, "FY2"))

totalapp_by_year <- rbind(app_00_06, app_07_15, app_16_22)

totalapp_by_year <- str_replace_all(totalapp_by_year$`U.S. Border Patrol Nationwide Apprehensions `, "[FY]", "") %>% 
  data.frame() 
colnames(totalapp_by_year)[colnames(totalapp_by_year) == "."] <- "year"
# totalapp_by_year
unique(totalapp_by_year)

apps_by_yr <- totalapp_by_year %>% 
  select(year) %>% 
  na.omit() %>% 
  count(year)

colnames(apps_by_yr)[colnames(apps_by_yr) == "n"] <- "Apprehensions at the Border"
apps_by_yr <- apps_by_yr %>% 
  transform(year = as.double(year))
# works
# as.numeric(apps_by_yr$year)
# works, but need to put into dataframe 


apps_by_yr
```


## 2 Election results 2000-2020 

```{r}
election_results <- read_delim("../../data_cleaned/harvard_election/countypres_2000-2020.csv")

election_results 

# election_results %>% 
#   select(year, party, candidatevotes) %>% 
#   summarize(year, party, candidatevotes = sum(candidatevotes)) %>% 
#   ggplot(aes(x = year, y = candidatevotes, fill = factor(party))) +
#   geom_col(stat = "identity", position = "dodge")

# original
election_results %>% 
  select(year, party, candidatevotes) %>% 
  summarize(year, party, candidatevotes) %>% 
  ggplot(aes(x = year, y = candidatevotes)) +
  # literally it was just getting rid of = sum(candidatevotes) from candidatevotes
  geom_col(aes(fill = party), stat = "identity", position = "dodge") +
  labs(title = "Election Results from Total Party Candidate Votes", 
       y = "Number of Candidate Votes")

# change linear x axis to categorical years?

# TODO: Make graph of # of states which were republican or democrat for each year

```
      # Dont think I need this
      
```{r}
# votestimeframe <- election_results %>% 
#   select(year, state, county_name, party, office, candidatevotes, totalvotes) %>%
#   filter(office == "US PRESIDENT") %>% 
#   group_by(year, state, county_name, party, office) %>% 
#   summarise(year, state, county_name, candidatevotes) %>%
#   ungroup() %>%
#   distinct(party, state, county_name, year, .keep_all = TRUE) %>%
#   select(party, state, county_name, year, candidatevotes) %>%
#   pivot_wider(id_cols = c(state, county_name), names_from = c(party, year), values_from = c(candidatevotes))
#   
# votestimeframe %>% 
#   head(10)
```

```{r}
date_sep <- function(data, date_col) {
  mutate(data, date = as_date({{date_col}}, format = "%m/%d/%Y")) %>%
  mutate(year = as.numeric(format(date, format = "%Y")),
         month = as.numeric(format(date, format = "%m")),
         day = as.numeric(format(date, format = "%d"))) %>% 
    select(year, month, day, date) %>% 
    na.omit()
}

date_sep(ice_rem_raw, ...2)

totalice_yr11_12_test <- ice_rem_raw %>%
  select(...2) %>%
  date_sep(...2)

totalice_yr11_12_test
```

## 3 Number of deportations (by year/total)

```{r}
# totaldep_by_year <- test_df %>% 
#   select(...8) %>% 
#   mutate(year = ...8 %in% c())
# totaldep_by_year

totalice_yr11_12 <- ice_rem_raw %>% 
  select(...2) %>%
  date_sep(...2)

totalice_yr13 <- ice_rem_raw2 %>% 
  select(`ERO-LESA Statistical Tracking Unit`) %>% 
  date_sep(`ERO-LESA Statistical Tracking Unit`) 

totalice_yr14_15 <- ice_rem_raw2015 %>% 
  select(...3) %>% 
  date_sep(...3)
 
totalice_yr16_14 <- ice_rem_raw16_14 %>% 
  select(`ERO-LESA Statistical Tracking Unit`) %>% 
  date_sep(`ERO-LESA Statistical Tracking Unit`) 

totalice_yr19_17 <- ice_rem_raw19_17 %>% 
  select(`ERO-LESA Statistical Tracking Unit`) %>% 
  date_sep(`ERO-LESA Statistical Tracking Unit`) 

totalice_yr23_20 <- ice_rem_raw23_20 %>% 
  select(`ERO-LESA Statistical Tracking Unit`) %>% 
  date_sep(`ERO-LESA Statistical Tracking Unit`)

# totalice_yr11_12
# totalice_yr13
# totalice_yr14_15
# totalice_yr16_14
# totalice_yr19_17

totalice_by_year <- rbind(totalice_yr11_12, totalice_yr13, totalice_yr14_15, totalice_yr16_14, totalice_yr19_17, totalice_yr23_20) %>% 
  select(date, year, month, day)
totalice_by_year
sample_n(totalice_by_year, 20)



icedeports_by_yr <- totalice_by_year %>% 
  select(year) %>% 
  na.omit() %>% 
  count(year)



colnames(icedeports_by_yr)[colnames(icedeports_by_yr) == "n"] <- "Deportations"
icedeports_by_yr
```

## 3 Deportations vs Election Results

        # In wrong chunk
```{r}
# vote_diff_by_yr <- wide_votes %>% 
#   mutate("Vote Difference (Republican minus Democrat)" = REPUBLICAN - DEMOCRAT)
# vote_diff_by_yr
# # works but should probably clarify which group is negative/being subtracted

```
Quick check online shows that these election results are fairly accurate/consistent with actual numbers

```{r}
dem_votes <- election_results %>% 
  select(year, party, candidatevotes) %>% 
  filter(party %in% "DEMOCRAT") %>% 
  group_by(year, party) %>% 
  summarize(tot_votes = sum(candidatevotes))
dem_votes

rep_votes <- election_results %>% 
  select(year, party, candidatevotes) %>% 
  filter(party %in% "REPUBLICAN") %>% 
  group_by(year, party) %>% 
  summarize(tot_votes = sum(candidatevotes))
rep_votes


border_states <- c("MAINE", "NEW HAMPSHIRE", "VERMONT", "NEW YORK", "PENNSYLVANIA", "OHIO", "MICHIGAN", "MINNESOTA", "NORTH DAKOTA", "MONTANA", "IDAHO", "WASHINGTON", "ALASKA")

bor_status <- election_results %>% 
  mutate(border_state = state %in% border_states) 
bor_status

dem_bor_votes <- bor_status %>% 
  select(year, party, candidatevotes, border_state) %>% 
  filter(party %in% "DEMOCRAT") %>% 
  group_by(year, party, border_state) %>% 
  summarize(tot_votes = sum(candidatevotes))
dem_bor_votes

rep_bor_votes <- bor_status %>% 
  select(year, party, candidatevotes, border_state) %>% 
  filter(party %in% "REPUBLICAN") %>% 
  group_by(year, party, border_state) %>% 
  summarize(tot_votes = sum(candidatevotes))
rep_bor_votes


total_votes_by_yr <- rbind(dem_votes, rep_votes) 
total_votes_by_yr

wide_votes <- total_votes_by_yr %>% 
  pivot_wider(id_cols = year, names_from = party, values_from = tot_votes)
wide_votes

total_bor_votes_by_yr <- rbind(dem_bor_votes, rep_bor_votes) 
total_bor_votes_by_yr

wide_bor_votes <- total_bor_votes_by_yr %>% 
  pivot_wider(id_cols = year, names_from = c(party, border_state), values_from = tot_votes)
  # k changing names_from worked but need to change col names
colnames(wide_bor_votes)[colnames(wide_bor_votes) == "DEMOCRAT_FALSE"] <- "Non-Border: Democrat"
colnames(wide_bor_votes)[colnames(wide_bor_votes) == "DEMOCRAT_TRUE"] <- "Border: Democrat"
colnames(wide_bor_votes)[colnames(wide_bor_votes) == "REPUBLICAN_FALSE"] <- "Non-Border: Republican"
colnames(wide_bor_votes)[colnames(wide_bor_votes) == "REPUBLICAN_TRUE"] <- "Border: Republican"
wide_bor_votes
  


long_df_bor <- wide_bor_votes %>% 
  pivot_longer(cols = c("Non-Border: Democrat", "Border: Democrat", "Non-Border: Republican", "Border: Republican"),
                 names_to = "Number of State Votes:", names_transform = list(n = as.double),
                 values_to = "number") %>% 
  # TODO: need to change var names to Dem: bor/Dem: non-bor or smth like that
  na.omit()
long_df_bor



vote_diff_by_yr <- wide_votes %>% 
  mutate("Vote Difference (Republican minus Democrat)" = REPUBLICAN - DEMOCRAT)
vote_diff_by_yr

app_vs_votes <- rbind(wide_votes, apps_by_yr, vote_diff_by_yr) %>% 
  arrange(year)
app_vs_votes

dep_vs_votes <- rbind(wide_votes, icedeports_by_yr, vote_diff_by_yr) %>% 
  arrange(year)
dep_vs_votes

app_vs_dep <- merge(apps_by_yr, icedeports_by_yr) %>% 
  # ok merge worked but theres a cutoff of certain years
  arrange(year)
# currently have a variable type issue
# change numeric to dbl
app_vs_dep

# # checking if 1st column was same for both
# dep_col <- names(icedeports_by_yr)
# app_col <- names(apps_by_yr)
# 
# dep_col[1] == app_col[1]
# dep_col[1] 
# app_col[1]

long_df_app <- app_vs_votes %>% 
    pivot_longer(cols = c("DEMOCRAT", "REPUBLICAN", "Apprehensions at the Border",
                          "Vote Difference (Republican minus Democrat)"),
                 names_to = "Number of:", names_transform = list(n = as.double),
                 values_to = "number") %>% 
  na.omit()
long_df_app

long_df_dep <- dep_vs_votes %>% 
    pivot_longer(cols = c("DEMOCRAT", "REPUBLICAN", "Deportations",
                          "Vote Difference (Republican minus Democrat)"),
                 names_to = "Number of:", names_transform = list(n = as.double),
                 values_to = "number") %>% 
  na.omit()
long_df_dep

long_df_dep_app <- app_vs_dep %>% 
    pivot_longer(cols = c("Apprehensions at the Border", "Deportations"),
                 names_to = "Number of:", names_transform = list(n = as.double),
                 values_to = "number") %>% 
  na.omit()
long_df_dep_app
```

```{r}
# convert units to hundreds of thousands or millions for better sense of scale

# Deportations vs Votes
long_df_dep %>% 
  ggplot(aes(x = year, y = number, col = `Number of:`)) +
  geom_point() +
  geom_line() +
  labs(title = "Deportations and Votes, by Year")

# Apprehensions vs Votes
long_df_app %>%
  ggplot(aes(x = year, y = number, col = `Number of:`)) +
  geom_point() +
  geom_line() +
  labs(title = "Apprehensions and Votes, by Year")
# 
# # Deportations Only
# long_df_dep %>% 
#   filter(`Number of:` %in% c("Deportations", "Vote Difference (Republican - Democrat)")) %>% 
#   ggplot(aes(x = year, y = number)) +
#   geom_point() +
#   geom_line() +
#   labs(title = "Deportations by Year")
# percentage inc/dec or compared to average # of deportations?
# add on rather than separate graph

# Apprehensions Only
apps_by_yr %>% 
  ggplot(aes(x = year, y = `Apprehensions at the Border`, group = 1)) +
  geom_point() +
  geom_line() +
  labs(title = "Apprehensions at the Border by Year")

# Vote Difference Only
vote_diff_by_yr %>% 
  ggplot(aes(x = year, y = `Vote Difference (Republican minus Democrat)`,
             color = `Vote Difference (Republican minus Democrat)`)) +
  geom_point(size = 2) +
  geom_line() + 
  geom_hline(yintercept = 1.5, color = "black") +
  labs(title = "Voting Difference (Rep. - Dem.), by Year") +
  scale_color_gradient(
    low = "blue",
    high = "red")
# not working bc need    col = [insert variable] in ggplot(aes())

# Border vs Non-Border States Votes
long_df_bor %>% 
  ggplot(aes(x = year, y = number, col = `Number of State Votes:`)) +
  # , fill = border_state
  geom_point() +
  geom_line() +
  labs(title = "Border vs Non-Border State Votes, by Year")

# TODO: Bor vs Non-Bor voting trends?
# e.g. By Main Candidates 


# Apprehensions vs Deportations
long_df_dep_app %>% 
  ggplot(aes(x = year, y = number, col = `Number of:`)) +
  geom_point() +
  geom_line() +
  labs(title = "Apprehensions at the Border and Deportations, by Year")

# slight conflict in name, variables named app but called Apprehensions 


# add color separations by 4 year periods btwn elections?
# TODO: separate points between years on Dems vs Rep won the election
```

